\section{Discussion}

\begin{frame}{When does the rule break?}
  \begin{itemize}
    \item \textbf{It's a threshold, not a gradual decline.} The model absorbs noise
      fine up to $10\%$, then snaps at $50\%$. Similar to how a child's grammar
      survives occasional speech errors around them.
    \item \textbf{The attractor gap catches it early.} It climbs from $0.07$ to $0.20$
      even while overall accuracy still looks fine.
    \item \textbf{Low noise delays learning, high noise prevents it.} At low noise
      the model still gets there eventually (the curves converge). At $50\%$ it
      plateaus and never learns the rule.
  \end{itemize}
\end{frame}

\begin{frame}{Does structure buy robustness?}
  \begin{itemize}
    \item The big question: does a model with built-in syntax (RNNG) hold up
      better than a plain LSTM? RNNG training is still in progress.
    \item \textbf{What we know from the LSTM:} it picks up the hierarchical rule
      from clean data, but drifts toward the nearest-noun shortcut as noise grows.
    \item If RNNG keeps the hierarchical rule where the LSTM gave up, that
      points to structure helping with robustness.
    \item If both models break at the same point, then maybe it's really about
      the quality of the input, not the architecture. That ties back to the
      poverty-of-the-stimulus question.
  \end{itemize}
\end{frame}

\begin{frame}{Takeaways and future work}
  \textbf{What we found}
  \begin{itemize}
    \item LSTMs can handle moderate SVA noise, but there's a sharp breaking point.
    \item Noise doesn't just make the model worse overall. It changes \emph{which rule}
      the model learns, pushing it toward the linear nearest-noun shortcut.
    \item At $50\%$ noise, more training actually makes things worse, not better.
  \end{itemize}
  \vspace{0.3cm}
  \textbf{Limitations and next steps}
  \begin{itemize}
    \item Our noise is synthetic and uniform; the corpus is templatic; trajectories
      are single-seed.
    \item Finish the RNNG comparison; try ON-LSTM as a middle ground.
    \item Future: more naturalistic noise patterns and larger-scale training.
  \end{itemize}
\end{frame}
