\section{Methods}
\label{sec:methods}

% OWNER: Jackson Wilke (methods/setup: data, noise, architectures per proposal)

The design isolates a single comparison: vary only the agreement noise in the
training data, holding the test set, vocabulary, training-set size, and
architecture fixed, and measure whether a model learns the hierarchical
agreement rule (agree with the true subject) or backs off to the linear
nearest-noun rule.

\subsection{Data and Noise Injection}
\label{sec:data}
We build on the subject--verb agreement (SVA) corpus of \citet{linzen2016},
\texttt{agr\_50\_mostcommon\_10K} ($\approx$1.58M English Wikipedia sentences,
with rare tokens anonymized so the vocabulary is capped at the 10K most frequent
types). Each sentence is annotated with the subject, the inflected verb, the
verb's position, and the number of \emph{intervening} nouns that disagree in
number with the subject---the ``attractors'' central to the
hierarchical-vs-linear distinction. We train on the real
(\texttt{orig\_sentence}) word forms.

To study robustness to noise we construct five training corpora that are
identical except for the rate at which SVA agreement is corrupted. For each
sentence, with probability $p$ we flip the verb's grammatical number
(singular\,$\leftrightarrow$\,plural), producing an ungrammatical agreement
relation while leaving every other lexical and syntactic property unchanged. The
flip is a deterministic inflection (\texttt{lemminflect}, with hand-specified
irregulars for high-frequency copulas and auxiliaries such as
\textit{is}/\textit{are}, \textit{was}/\textit{were}, \textit{has}/\textit{have})
applied at the annotated verb index; the \emph{same} function builds the
ungrammatical form at evaluation time, so corruption and evaluation are
guaranteed consistent. The five noise rates are
$p \in \{0, 0.004, 0.02, 0.10, 0.50\}$ (baseline, low, medium-low, medium-high,
high). We verified each corpus by confirming that the empirical fraction of
corrupted sentences matches $p$ and that the baseline corpus is fully
grammatical. \textbf{Training-set size is held constant} across conditions, so
any change in behavior is attributable to noise rather than data quantity.

\subsection{Leak-Free Evaluation Protocol}
\label{sec:leakfree}
Because the five corpora are row-aligned with the source corpus, a row index
denotes the same sentence in every condition. We reserve a single fixed set of
20{,}000 held-out row indices, drawn once with a fixed seed (1234), as the test
set. These rows are \emph{excluded from every training corpus}, so no test
sentence is ever seen in training under any noise condition. The test items are
built from the \emph{uncorrupted} source corpus at those same indices: all test
sentences are therefore grammatical, and the identical test set scores every
condition. Only the training noise varies.

\subsection{Minimal Pairs}
\label{sec:minimalpairs}
For each held-out sentence we form a minimal pair: the \emph{prefix} is the
sentence up to (but not including) the verb; the \emph{correct} continuation is
the original grammatical verb $v^{+}$; the \emph{incorrect} continuation is its
number-flipped form $v^{-}$. A model is scored correct iff it assigns higher
conditional probability to the grammatical form,
\[
  P(v^{+}\mid \text{prefix}) > P(v^{-}\mid \text{prefix}),
\]
a single-next-token comparison in the style of BLiMP/SyntaxGym minimal-pair
evaluation. We keep only pairs where both verb forms are single in-vocabulary
tokens, so the comparison is well defined; this yields \textbf{19{,}819 usable
pairs (1{,}528 with an attractor, 18{,}291 without)}, identical across all
conditions. Each pair is tagged \texttt{has\_attractor} when at least one
intervening noun mismatches the subject in number.

\subsection{Models}
\label{sec:models}
We compare two architectures trained from scratch on each noise condition. The
primary contrast is between a \emph{sequential} model and a
\emph{syntactically-biased} model, directly targeting our question of whether an
explicit structural bias makes hierarchical agreement more robust to noisy
input.

\paragraph{Sequential baseline (LSTM).} A word-level LSTM language model in the
style of \citet{linzen2016} and \citet{gulordava2018}: two layers, tied
input/output embeddings, and dropout. Each sentence is modeled independently (no
streaming across boundaries).

\paragraph{Syntactically-biased model (RNNG).} A Recurrent Neural Network
Grammar \citep{dyer2016rnng}, which jointly models structure and terminals
through a generative sequence of \textsc{shift}/\textsc{reduce}/\textsc{nt}($X$)
actions and has been shown to improve syntactic generalization over sequential
models \citep{wilcox2020}. We use the \texttt{rnng-pytorch} implementation
\citep{noji2021}. RNNG training requires bracketed parse trees; we obtain them
once with the Berkeley Neural Parser \citep{kitaev2018}. Crucially, the verb
flip changes only the verb \emph{terminal}, not the constituency bracketing, so
the parse is identical across all five noise conditions: we parse the clean
corpus once and swap the verb terminal per condition, making tree generation a
one-time cost rather than $5\times$. We had prepared ON-LSTM \citep{shen2019}, a
softer structural bias requiring no gold trees, as a fallback; it was not needed.

\subsection{Metrics and Rule Probing}
\label{sec:metrics}
Our primary metric is minimal-pair accuracy (\texttt{acc\_all}); as a graded
secondary measure we report the surprisal $-\log_2 P(v^{+}\mid\text{prefix})$ of
the correct verb. To read off \emph{which} rule a model learned, we split
accuracy by attractor presence and report the \textbf{attractor gap}
\[
  \Delta = \mathrm{acc}(\text{no-attractor}) - \mathrm{acc}(\text{attractor}).
\]
A small $\Delta$ indicates the hierarchical rule (attractors do not hurt); a
large $\Delta$ indicates the linear nearest-noun rule. Our central analysis is
$\Delta$ as a function of the training-noise rate.

\paragraph{Cross-model comparability.} The LSTM scores all 19{,}819 in-vocab
pairs directly. The RNNG is scored by word-synchronous beam search, which yields
per-verb surprisals for only the subset of held-out sentences it can parse
within the beam (and does not apply the in-vocabulary verb filter). To keep the
LSTM-vs-RNNG contrast on identical items, we evaluate both models on the
\emph{matched subset}: the intersection of the pairs the RNNG can score with the
LSTM's in-vocabulary pairs. The headline comparison and the H3 test
(Section~\ref{sec:experiments}) are computed on this matched subset; full
reproduction details are in Appendix~\ref{app:repro}.
