\section{Conclusion}
\label{sec:conclusion}

This study investigates how grammatical noise affects the acquisition of grammatical rules in language models by systematically injecting subject--verb agreement (SVA) errors into the training data and comparing the robustness of models with different structural inductive biases. By varying only the level of grammatical corruption while holding the test set, vocabulary, dataset size, and model configuration constant, we are able to directly observe how grammatical knowledge is learned under degraded linguistic input.

Our results show that language models exhibit a substantial degree of grammatical robustness. Under low to moderate levels of noise, models are still able to acquire the correct subject--verb agreement rule, suggesting that hierarchical grammatical knowledge does not require perfectly clean input to emerge. However, this robustness is not unlimited. Once the noise level becomes sufficiently high, performance deteriorates sharply and exhibits a clear threshold effect: rule acquisition is no longer merely delayed, but effectively prevented.

Further analysis reveals that noise influences not only overall performance but also the type of generalization strategy adopted by the model. As the proportion of grammatical errors increases, the attractor gap of the LSTM grows steadily, indicating a gradual shift from a hierarchical agreement strategy based on the true subject toward a simpler linear nearest-noun heuristic. In other words, grammatical noise reshapes the model's internal syntactic representations and encourages reliance on a simpler but less robust form of generalization.

Comparing model architectures further shows that structural inductive bias improves robustness to noisy input. Across all noise conditions, the RNNG consistently outperforms the LSTM and remains substantially more stable under heavy corruption. In particular, the RNNG maintains a much smaller attractor gap, indicating that it preserves hierarchical syntactic generalization even when grammatical evidence becomes unreliable. These findings suggest that structural inductive bias influences not only learning efficiency but also the robustness of grammatical rule acquisition.

From a broader cognitive-science perspective, our findings provide new computational evidence relevant to the Poverty of the Stimulus debate. On the one hand, the ability of language models to acquire correct grammatical rules under substantial noise suggests that hierarchical structure may be more learnable from input than strong nativist accounts would predict. On the other hand, the increasing reliance on linear heuristics under heavy noise, together with the superior robustness of the structurally biased RNNG, indicates that internal inductive biases play an important role in the acquisition and maintenance of complex grammatical knowledge. Taken together, our results support neither a purely empiricist nor a strongly nativist account of language acquisition. Instead, they point toward an interactionist view in which grammatical knowledge emerges through the combined influence of input statistics and structural inductive biases.

Future work could extend this framework to more naturalistic distributions of grammatical errors, larger-scale language models, and a broader range of syntactic phenomena, providing further insight into how language models learn grammar under realistic noisy conditions and what this may reveal about human language acquisition.

% OWNER: Kelly Fu (intro/conclusion per proposal)
% TODO (Kelly): Restate question + setup in 1--2 sentences; headline finding;
%       future work.
